Parsing Postfix log files

### 2008-08-01 doc/logparser-architecture.tex
Improved parts of the efficiency section and inlined graphs. Fixed up some references; added $Id$ to front page.
Bunch of XXX items added.
Dealt with some XXX items.

@@

    \item [actions] Actions are invoked to deal with a log line once it has
        been identified by the rules: actions modify data structures,
        handle complications, and cause data to be saved to the database.
        The actions perform the work of reconstructing the journey a mail
        takes through Postfix.  Full details of actions can be found in
        \sectionref{actions-in-detail} and \sectionref{adding new actions}
        XXX IMPROVE THIS SENTENCE\@.

    \item [framework] The framework is responsible for loading rules,
        managing data structures, finding the rule which matches each log
        line, invoking the correct action, etc\@.  The framework provides
        the structure which actions and rules plug into.  The framework is
        described in detail in \sectionref{framework}.  XXX EXTEND
        FRAMEWORK SECTION IF NECESSARY\@.

@@

XXX CLARIFY THE SENTENCE STARTING WITH ``Although this may see'' IN THE
NEXT PARAGRAPH\@.

Separating the rules from the actions and framework makes it possible to
parse new log lines without modifying the core parsing algorithm.  Although
this may seem like a trivial point, is it substantially more difficult to
understand a program's entire parsing algorithm, identify the correct
location to change and make the appropriate changes, versus adding a new
rule with the action to invoke, a \regex{} to match the log lines, and the
specification of the data to extract.  Bear in mind that the changes must
be made without adversely affecting existing parsing, particularly as there
may be edge cases which are not immediately obvious.\footnote{See
\sectionref{yet-more-aborted-delivery-attempts} for a complication which
occurs only four times in \numberOFlogFILES{} log files tested.}  Requiring
changes to the parser's code also complicates upgrades, as the changes must
be preserved during the upgrade, and may clash with changes made by the
developer.\footnote{See \sectionref{adding new actions} for how to add new
actions}  \parsername{} allows the user to add new rules to the database
without changing the parsing algorithm, unless the new log lines to be
parsed require functionality not already provided by the algorithm.  If the
new log lines do require new functionality, new actions can be added to the
parser without modifying existing actions or other parts of the algorithm;
only in the rare case that the new actions require support from other
sections of the code will more extensive changes be required.

XXX CHANGE THIS\@; THE PARSER DESIGN IS NOVEL, DO NOT STRESS THE SIMILARITY
TOO MUCH\@.

There is some similarity between the parser's design and William Wood's
\ATN{}~\cite{atns, nlpip}, a tool used in Computational Linguistics for
creating grammars to parse or generate sentences.  The resemblance between
\ATN{} and the parser is accidental, but it is interesting how two
apparently different approaches share an underlying separation of concerns;
this appears to be a natural division of responsibility and functionality.

@@

\begin{tabular}[]{lll}
    \textit{\ATN{}\/}   & \textit{Parser\/} & \textit{Similarity\/}     \\
    Networks            & Algorithm         & Determines the sequence 
                                              of transitions            \\
                        &                   & or actions which 
                                              constitutes a valid       \\
                        &                   & input.                    \\
    Transitions         & Actions           & Save data and impose
                                              conditions the            \\
                        &                   & input must meet to be
                                              considered valid.         \\
    Abbreviations       & Rules             & Responsible for 
                                              classifying input.        \\
\end{tabular}

@@

    \item Compare the results of parsing using sorted, shuffled and
        reversed rules (described in \sectionref{rule efficiency}).  Parse
        a number of log files using normal sorting, then dump a textual
        representation of the rules, connections and results tables.
        Repeat with shuffled and reversed rules, starting with a fresh
        database.  If there are no overlapping rules the tables from each
        run will be identical; differences indicate overlapping rules.
        Which rules overlap can be determined by examining the differences
        in the tables: each result contains a reference to the rule which
        created it, if the references differ between runs the two rules
        referenced in the differing results overlap.  Unfortunately this
        method cannot prove the absence of overlapping rules; it can detect
        overlapping rules, but only if there are log lines in the input
        files which match more than one rule.

@@

>>>>

XXX REWRITE AS NECESSARY TO MERGE AND SUPPLEMENT THE EXISTING CONTENT\@.

@@

    \item [actions] Actions are invoked to deal with a log line once it has
        been identified by the rules: actions modify data structures,
        handle complications, and cause data to be saved to the database.
        The actions perform the work of reconstructing the journey a mail
        takes through Postfix.  Details of the actions available in the
        Postfix parser can be found in \sectionref{actions-in-detail}, and
        \sectionref{adding new actions} describes the process of adding new
        actions.

    \item [framework] The framework is responsible for loading rules,
        managing data structures, finding the rule which matches each log
        line, invoking the correct action, etc\@.  The framework is
        described in detail in \sectionref{framework}.  XXX EXTEND
        FRAMEWORK SECTION IF NECESSARY\@.

@@

Separating the rules from the actions and framework makes it possible to
parse new log lines without modifying the core parsing algorithm.  Adding a
new rule with the action to invoke and a \regex{} to match the log lines is
trivial in comparison to understanding a program's entire parsing
algorithm, identifying the correct location to change and making the
appropriate changes.  Bear in mind that the changes must be made without
adversely affecting existing parsing, particularly as there may be edge
cases which are not immediately obvious ---
\sectionref{yet-more-aborted-delivery-attempts} describes a complication which
occurs only four times in \numberOFlogFILES{} log files tested.  Requiring
changes to the parser's code also complicates upgrades, as the changes must
be preserved during the upgrade, and may clash with changes made by the
developer (\sectionref{adding new actions} describes safely adding new
actions).  \parsername{} allows the user to add new rules to the database
without changing the parsing algorithm, unless the new log lines to be
parsed require functionality not already provided by the algorithm.  If the
new log lines do require new functionality, new actions can be added to the
parser without modifying existing actions or other parts of the algorithm;
only in the rare case that the new actions require support from other
sections of the code will more extensive changes be required.

There is some similarity between the parser's design and William Wood's
\ATN{}~\cite{atns, nlpip}, a tool used in Computational Linguistics for
creating grammars to parse or generate sentences.  The resemblance between
\ATN{} and the parser is accidental, but it is interesting how two
apparently different approaches share an underlying separation of concerns;
this appears to be a natural division of responsibility and functionality.
\Tableref{Similarities between ATN and this architecture} shows the
similarities.

@@

\begin{table}[ht]
    \caption{Similarities between ATN and this architecture}
    \empty{}\label{Similarities between ATN and this architecture}
    \begin{tabular}[]{lll}
        \tableline{}
        \ATN{}          & Architecture  & Similarity                \\
        \tableline{}
        Networks        & Algorithm     & Determines the sequence 
                                        of transitions              \\
                        &               & or actions which 
                                        constitutes a valid  input. \\
        Transitions     & Actions       & Save data and impose
                                        conditions the              \\
                        &               & input must meet to be
                                        considered valid.           \\
        Abbreviations   & Rules         & Responsible for 
                                        classifying input.          \\
        \tableline{}
    \end{tabular}
\end{table}

@@

    \item Compare the results of parsing using sorted, shuffled and
        reversed rules (described in \sectionref{rule ordering for
        efficiency}).  Parse a number of log files using normal sorting,
        then dump a textual representation of the rules, connections and
        results tables.  Repeat with shuffled and reversed rules, starting
        with a fresh database.  If there are no overlapping rules the
        tables from each run will be identical; differences indicate
        overlapping rules.  Which rules overlap can be determined by
        examining the differences in the tables: each result contains a
        reference to the rule which created it, if the references differ
        between runs the two rules referenced in the differing results
        overlap.  Unfortunately this method cannot prove the absence of
        overlapping rules; it can detect overlapping rules, but only if
        there are log lines in the input files which match more than one
        rule.

@@

\subsection{Actions}

XXX TO BE WRITTEN\@.

\subsection{Rules}

XXX TO BE WRITTEN\@.

### 2008-08-01 doc/logparser-background.tex
Wrote minimum ruleset versus maximum ruleset section - largely taken from the SGAI paper.
Dealt with some XXX items.

Each restriction is evaluated to produce a result of \textit{reject},
\textit{permit}, \textit{dunno\/} or the name of another restriction to be
evaluated.\footnote{Other results are possible as described
in~\cite{smtpd_access_readme, smtpd_per_user_control, policy-servers}.} The
meaning of \textit{permit\/} and \textit{reject\/} is fairly obvious;
\textit{dunno\/} means to stop evaluating the current restriction and
continue processing the remainder of the restriction list, allowing more
specific cases to be used as exceptions to more general cases.  When the
result is the name of another restriction Postfix will evaluate the new
restriction, allowing restrictions to be chosen based on the client \IP{}
address, HELO hostname, sender address, recipient address,
etc.\footnote{E.g.\ the administrator may require that all clients on the
local network have valid \DNS{} entries, to prevent people sending mail
from unknown machines.}  The administrator can define new restrictions as a
list of existing restrictions, allowing arbitrarily long and complex
sequences of lookups, restrictions and exceptions.  Postfix tries to
protect the administrator in as far as is reasonable, e.g.\ the restriction
\texttt{check\_helo\_mx\_access} cannot cause a mail to be accepted,
because the parameter it checks (the hostname given in the HELO command) is
under the control of the remote client.  Despite this, it is possible for
the administrator to make catastrophic mistakes, e.g.\ rejecting mail for
all users --- the administrator must be cognisant of the ramifications of
their configuration changes.  XXX UNIX DOESN'T STOP YOU DOING STUPID THINGS
BECAUSE THAT WOULD ALSO STOP YOU DOING CLEVER THINGS\@.

>>>>

Each restriction is evaluated to produce a result of \textit{reject},
\textit{permit}, \textit{dunno\/} or the name of another restriction to be
evaluated.\footnote{Other results are possible as described
in~\cite{smtpd_access_readme, smtpd_per_user_control, policy-servers}.} The
meaning of \textit{permit\/} and \textit{reject\/} is fairly obvious;
\textit{dunno\/} means to stop evaluating the current restriction and
continue processing the remainder of the restriction list, allowing more
specific cases to be used as exceptions to more general cases.  When the
result is the name of another restriction Postfix will evaluate the new
restriction, allowing restrictions to be chosen based on the client \IP{}
address, HELO hostname, sender address, recipient address,
etc.\footnote{E.g.\ the administrator may require that all clients on the
local network have valid \DNS{} entries, to prevent people sending mail
from unknown machines.}  The administrator can define new restrictions as a
list of existing restrictions, allowing arbitrarily long and complex
sequences of lookups, restrictions and exceptions.  Postfix tries to
protect the administrator in as far as is reasonable, e.g.\ the restriction
\texttt{check\_helo\_mx\_access} cannot cause a mail to be accepted,
because the parameter it checks (the hostname given in the HELO command) is
under the control of the remote client.  Despite this, it is possible for
the administrator to make catastrophic mistakes, e.g.\ rejecting mail for
all users --- the administrator must be cognisant of the effects of their
configuration changes.  This is similar to one of UNIX's design
philosophies: \textit{``UNIX was not designed to stop its users from doing
stupid things, as that would also stop them from doing clever things''\/}
--- Doug Gwyn~\cite{unix-philosophy}.

### 2008-08-01 doc/logparser.tex
Lots of improvements to the efficiency section - I'm happy with most of it now. I improved the graphs too, with better captions and labels.
Wrote minimum ruleset versus maximum ruleset section - largely taken from the SGAI paper.
Further corrections from Carl. Added paragraph about pathological rules. Some cleanups and XXX items dealt with.
Removed the graphs showing breakdown of rule hits. Moved remaining graphs out of graphs section and removed graphs section. Minor corrections.
Dealt with some XXX items.

@@

XXX CHECK Y-AXIS LABELS ON GRAPHS TO ENSURE THEY CAN BE READ, AND WRAP IF
NECESSARY\@.

XXX EXPLAIN THE X-AXIS ON GRAPHS --- MADS THOUGHT IT WAS UNCLEAR\@.

XXX CHECK THE CAPTIONS ON ALL GRAPHS AND TABLES --- SOME ARE AWFUL\@.

>>>>

\newcommand{\numberOFrulesMINIMUM}[0]{%
    115%
}

% \numberOFrulesMINIMUM as percentage of \numberOFrules
\newcommand{\numberOFrulesMINIMUMpercentage}[0]{%
    68.04\%%
}

% \numberOFrules as percentage increase of \numberOFrulesMINIMUM
\newcommand{\numberOFrulesMAXIMUMpercentage}[0]{%
    46.95\%%
}

@@

XXX GENERATE RESULT\_COLS AND CONNECTION\_COLS AUTOMATICALLY\@.

XXX NEED A NAME FOR THE ARCHITECTURE, AND A DIFFERENT NAME FOR THE
IMPLEMENTATION\@.
